---
title: "HW 5"
author: "Ian Douglas"
date: "3/11/2019"
output: word_document
---
#HW 5
####Read in data
```{r}
df <- read.table("~/Desktop/Grad School/Columbia/Spring 2019/Computational Stats/R/HW/Datasets/ebay.txt",header=T)
#Define variables
sp5 <- as.vector(df$SP500)
ebay <- as.vector(df$EBAY)
```
##Part A
$$Return_{Day_{i}} = \frac{Price_{Day_{i}}-Price_{Day_{i-1}}}{Price_{Day_{i-1}}}$$
```{r}
#Since the data is in chronologically descending order, index i PLUS 1 is
#actually day i MINUS 1
rtrn.sp5 <- NULL
rtrn.ebay <- NULL
for (i in 1:(nrow(df)-1)) {
  rtrn.sp5[i] <- ((sp5)[i] - (sp5[i+1]))/(sp5[i+1])
  rtrn.ebay[i] <- ((ebay)[i] - (ebay[i+1]))/(ebay[i+1])
}
returns <- cbind(matrix(rtrn.sp5,ncol=1),matrix(rtrn.ebay,ncol=1))
colnames(returns) <- c("S&P 500", "Ebay Stock")
rownames(returns) <- df$DATE[1:60]
returns
```
   
   ##Part B
```{r}
#Correlation between the two lists of returns
cor(returns[,1],returns[,2])
```
   
####Output shows a correlation of .3094   
   
   ##Part C
```{r}
#Estimate the SE of the correlation coefficient using 1000 bootstrap samples
library("boot")
cor.boot <- function(data, i){
	dat <- data
  dat2 <- dat[i,]
  cor.boot <- cor(dat2[i,1],dat[i,2])
	return(cor(dat2[,1], dat2[,2]))
}

corrs <- boot(returns, cor.boot, R=1000)
SE <- sd(corrs$t[,1])/sqrt(corrs$R)
SE
```
   
   ####Output shows a SE of .003

##Part D
```{r}
# Report the BCa 95% CI for the correlation coefficient:
boot.ci(corrs, type="all")
```
   
####Output shows the 95% BCa CI to be (.0473, .4705)

##Part E
Calculate $z_r±1.96\sqrt{\frac{1}{n-3}}$ where $z_r = \frac{1}{2}log\sqrt{\frac{1+r}{1-r}}$
```{r}
zr <- NULL
for (i in 1:corrs$R) {
  zr[i] <- (.5)*log(sqrt((1+mean(corrs$t[i,1]))/(1-mean(corrs$t[i,1]))))
}
zr.mean <- mean(zr)
int <- 1.96*sqrt(1/(nrow(returns)-3))
Fisher.ci <- c(zr.mean-int,zr.mean+int)
Fisher.ci
```
   
   ####Output shows a CI from -0.096 to .424
   
##Part F
```{r}
tanh(Fisher.ci)
# [1] -0.09528203  0.39999647
```
   
   ####Compared to the CI in part D, the re-scaled Fisher CI is larger.
   It also crosses zero, bringing into question the statistical significance
   of the correlation.
   more