---
title: "Day 2 Notes"
author: "Ian Douglas"
date: "2/5/2019"
output: html_document
---
#making user-defined functions
```{r}
#generic example
myfunction <- function(arg1, arg2, ...) {
  #this means we are making a function that takes arg1 and arg2, and
  #others, denoted by the elipses.
  
  #then define a statement
  sum = arg1 + arg2
  #then print by defining: return(object)
  return(sum)
  #note, you don't need to do return(object), you can just write `object`
}
```
#specific example
```{r}
sumsq <- function(vec) {
  mn <- mean(vec)
  d <- vec - mn
  d2<-d^2
  ss <- sum(d2)
  ss
}
```

###Definition
#Monte carlo study is to test a theory by simulating that a certain
#distribution will look a certain way, or a population of this distribution
#would have some given (theorized) variance or mean.
###Purpose:
#Simulate data from this theorized distribution and study its expected
#values, means, or if it comes from a known distribution, do new 
#estimators of known parameters have bias, large or small MSE, etc. How
#does this change with sample size, with non-constant variance, etc.
```{r}

```
